Asymmetric Agents and their Willingness to Fight\\ {\Large Customer Loyalty and its Impact on Competition}

### 2015-10-26 Paper1.tex
some suggestions

The optimal effort level for both firms and the indifferent customer are:
\begin{align}
\begin{split}
x_1^C &= g(x_1^C-x_2^C)\frac{\gamma_1}{c_1} + G(x_1-x_2)\frac{\partial \gamma_1}{\partial x_1}\frac{1}{c_1}\\
x_2^C &= g(x_1^C-x_2^C)\frac{\gamma_2}{c_2} + \big(1-G(x_1-x_2)\big)\frac{\partial \gamma_2}{\partial x_2}\frac{1}{c_2}
\end{split}
\end{align}
\begin{align}
\theta^* &=  g(x_1^C-x_2^C)\left(\frac{\gamma_1}{c_1}-\frac{\gamma_2}{c_2}\right) + G(x_1-x_2)\frac{\partial \gamma_1}{\partial x_1}\frac{1}{c_1} - \big(1-G(x_1-x_2)\big)\frac{\partial \gamma_2}{\partial x_2}\frac{1}{c_2}
\end{align}
As long as $g(\cdot)>0$, efforts are non-zero. For simplicity we normalize the costs of effort provision to $1$ without loss of generality, for both firms, by defining: $\gamma_i(x_i)\equiv \gamma_i(x_i)/c_i$. Depending on the shape of $G(\cdot)$ the game might feature multiple equilibria. To avoid having to assume uniqueness the following discussion focuses only on local changes.

@@

If $-\frac{1}{\gamma_1}<g'(x_1^C-x_2^C)<0$ this expression is positive, such that an increase in $x_2$ leads to an increase in $x_1$ and thus an overall escalation. The increase in $x_2$ raises the number of indifferent customers, such that competition becomes more intense. If $g'(x_1^C-x_2^C)>0$ this expression is negative and an increase in $x_2$ leads to crowding out of $x_1$. \marginpar{add a sketch here}

Returning to the example of the smartphone sector, we consider a distribution that satisfies $-\frac{1}{\gamma_1}<g'(\theta)<\frac{1}{\gamma_2}$\footnote{This assures us that the second order condition is satisfied for both firms.} for all $\theta$.\marginpar{I need to address symmetry earlier in the text.} Additionally, most customers are centered in the middle with only a few highly loyal customers. This can be described by $g'(0)=0$ and $g'(\theta)<0~\forall \theta>0$ and $g'(\theta)<0~\forall \theta<0$. Most customer are willing to switch to the superior product, but a small number has some inherent preference for one.

\subsection{General Case}
Now, we analyze the more general case with $\frac{\partial \gamma_1}{\partial x_1}\geq 0$. The reaction function of firm $1$ to a change in the efforts of firm $2$ is given by:
\begin{align}
\frac{\d x_1}{\d x_2} &= g'(\theta^*)\left(\tfrac{\d x_1}{\d x_2}-1\right)\gamma_1 + g(\theta^*)\left(2\tfrac{\d x_1}{\d x_2}-1\right)\tfrac{\partial \gamma_1}{\partial x_1} + G(\theta^*)\tfrac{\partial^2 \gamma_1}{\partial^2 x_1}\tfrac{\d x_1}{\d x_2}\nonumber\\
\frac{\d x_1}{\d x_2} &= A_i^{-1} \left(-g'(\theta^*)\gamma_1- g(\theta^*)\frac{\partial \gamma_1}{\partial x_1}\right)\\
A_i &= 1 - g'(\theta^*)c_1-2g(\theta^*)\frac{\partial \gamma_1}{\partial x_1}-G(\theta^*)\frac{\partial^2 \gamma_1}{\partial^2 x_1}
\end{align}
Similar to before, the second order condition implies that $A_i>0$ and by assumption $\frac{\partial \gamma_1}{\partial x_1}>0$. The sign of the reaction function depends on $g'(\cdot)$ and $g(\cdot)$, and $\frac{\partial \gamma_1}{\partial x_1}$ and $\gamma_i$. In fact the reaction depends on the change to the return of the indifferent customer. The reaction is positive if $\frac{\partial}{\partial x_1} g(\theta^*)\gamma_1<0$. The competition intensity described by $g(\theta^*)$ increases sufficiently to raise the incentives to compete more intensively. If $g'(\cdot)$ is close to zero, competition intensity does still increase, but is offset by the lower market share. A lower market share makes extracting profits from the customers less valuable, thus making $\frac{\partial \gamma_i}{\partial x_i}$ less important. Thus, for $\frac{\partial}{\partial x_1} g(\theta^*)\gamma_1>0$ and increase in $x_2$ leads to lower competition and a lower effort by $1$.

@@

@@

In contrast, if $\frac{\d x_1}{\d x_2}\ll0$ and/or if $\frac{\partial U(x_1,x_2)}{\partial x_1}\gg \frac{\partial U(x_1,x_2)}{\partial x_2}$ the third party should reduce the incentives of firm $2$ to reduces its efforts and encourage firm $1$ to increase its efforts. While the total amount of effects does necessarily decrease, firm $1$'s efforts are seen as more important than firm $2$, creating a net benefit. If the effort do not change the firm's profit extraction, the latter case requires that $\frac{1}{\gamma_1}<g'(\theta^*)<0$.

In the smartphone sector we observe: a small level of loyalty and a medium amount of asymmetry. Thus, supporting the weak firm will raise overall competition leading to a welfare increase. Similar, damaging the profit margin of the leader might lead to an overall increase in the efforts if the change in the number of indifferent customers is sufficiently big. Here moving closer to symmetry increases competition intensity.

@@

@@

@@

If customer loyalty is missing from the model, the only channel through which the efforts of the firms interact is the value extraction. If firm $2$ expands its market share at the expenses of firm $1$, firm $2$ has an additional incentive to have higher efforts and firm $1$ to reduce its efforts. This fact does not depend on the market shares, so supporting the weak firm always reduces the efforts of the leader.

@@

In the next period, both firms exert less efforts as $g(\theta^*)$ goes down. However, the stronger firm ($1$) reacts more strongly to this, shifting market shares to firm $2$. Ironically, the market shares become more equal. It is important to note, that this builds on myopic firms.  
\section{Other aspects}
\begin{itemize}
	\item Market shares determine investments. Are returns convex or concave?
	\item Customers profit from investment. The large firm has more customers. If it invests more, it helps more people. 
	\item What is the social optimum?
	\item Switching costs?
\end{itemize}

>>>>

The optimal effort level for both firms and the indifferent customer are:
\begin{align}
\begin{split}
x_1^C &= g(x_1^C-x_2^C)\frac{\gamma_1}{c_1} + G(x_1-x_2)\frac{\partial \gamma_1}{\partial x_1}\frac{1}{c_1}\\
x_2^C &= g(x_1^C-x_2^C)\frac{\gamma_2}{c_2} + \big(1-G(x_1-x_2)\big)\frac{\partial \gamma_2}{\partial x_2}\frac{1}{c_2}
\end{split}
\end{align}
\begin{align}
\theta^* &=  g(x_1^C-x_2^C)\left(\frac{\gamma_1}{c_1}-\frac{\gamma_2}{c_2}\right) + G(x_1-x_2)\frac{\partial \gamma_1}{\partial x_1}\frac{1}{c_1} - \big(1-G(x_1-x_2)\big)\frac{\partial \gamma_2}{\partial x_2}\frac{1}{c_2}
\end{align}

\marginpar{I am not sure I understand the third equation; how does this relate to $\theta^{*} = x_1 - x_2$???; what does the superscript $C$ mean?}

As long as $g(\cdot)>0$, efforts are non-zero. For simplicity we normalize the costs of effort provision to $1$ without loss of generality, for both firms, by defining: $\gamma_i(x_i)\equiv \gamma_i(x_i)/c_i$. Depending on the shape of $G(\cdot)$ the game might feature multiple equilibria. To avoid having to assume uniqueness the following discussion focuses only on local changes.

@@

If $-\frac{1}{\gamma_1}<g'(x_1^C-x_2^C)<0$ this expression is positive, such that an increase in $x_2$ leads to an increase in $x_1$ and thus an overall escalation. The increase in $x_2$ raises the number of indifferent customers, such that competition becomes more intense.\marginpar{explicitly say what your definition of competition intensity is: $dx_1/dx_2 >0$? or $d(x_{1}+x_2)/dx_2>0$ or increase in number of indifferent consumers? In the latter case, it should be something like marginal over infra-marginal?} If $g'(x_1^C-x_2^C)>0$ this expression is negative and an increase in $x_2$ leads to crowding out of $x_1$. \marginpar{add a sketch here}

Returning to the example of the smartphone sector, we consider a distribution that satisfies $-\frac{1}{\gamma_1}<g'(\theta)<\frac{1}{\gamma_2}$\footnote{This assures us that the second order condition is satisfied for both firms.} for all $\theta$.\marginpar{I need to address symmetry earlier in the text.} Additionally, most customers are centered in the middle with only a few highly loyal customers. This can be described by $g'(0)=0$ and $g'(\theta)<0~\forall \theta>0$ and $g'(\theta)<0~\forall \theta<0$. Most customer are willing to switch to the superior product, but a small number has some inherent preference for one.\marginpar{both $g'<0$?}

\subsection{General Case}
Now, we analyze the more general case with $\frac{\partial \gamma_1}{\partial x_1}\geq 0$. The reaction function of firm $1$ to a change in the efforts of firm $2$ is given by:
\begin{align}
\frac{\d x_1}{\d x_2} &= g'(\theta^*)\left(\tfrac{\d x_1}{\d x_2}-1\right)\gamma_1 + g(\theta^*)\left(2\tfrac{\d x_1}{\d x_2}-1\right)\tfrac{\partial \gamma_1}{\partial x_1} + G(\theta^*)\tfrac{\partial^2 \gamma_1}{\partial^2 x_1}\tfrac{\d x_1}{\d x_2}\nonumber\\
\frac{\d x_1}{\d x_2} &= A_i^{-1} \left(-g'(\theta^*)\gamma_1- g(\theta^*)\frac{\partial \gamma_1}{\partial x_1}\right)\\
A_i &= 1 - g'(\theta^*)c_1-2g(\theta^*)\frac{\partial \gamma_1}{\partial x_1}-G(\theta^*)\frac{\partial^2 \gamma_1}{\partial^2 x_1}
\end{align} \marginpar{why subscript $i$? $A_1$?}
Similar to before, the second order condition implies that $A_i>0$ and by assumption $\frac{\partial \gamma_1}{\partial x_1}>0$. The sign of the reaction function depends on $g'(\cdot)$ and $g(\cdot)$, and $\frac{\partial \gamma_1}{\partial x_1}$ and $\gamma_i$. In fact the reaction depends on the change to the return of the indifferent customer. The reaction is positive if $\frac{\partial}{\partial x_1} g(\theta^*)\gamma_1<0$. The competition intensity described by $g(\theta^*)$ increases sufficiently to raise the incentives to compete more intensively. If $g'(\cdot)$ is close to zero, competition intensity does still increase, but is offset by the lower market share. A lower market share makes extracting profits from the customers less valuable, thus making $\frac{\partial \gamma_i}{\partial x_i}$ less important. Thus, for $\frac{\partial}{\partial x_1} g(\theta^*)\gamma_1>0$ and increase in $x_2$ leads to lower competition and a lower effort by $1$.

@@

\marginpar{I would say: two reasons to raise $x1$: higher market share ($g$) and higher margin over existing market share. Relevant effect: how does $x_2$ affect these two margins?}

@@

In contrast, if $\frac{\d x_1}{\d x_2}\ll0$ and/or if $\frac{\partial U(x_1,x_2)}{\partial x_1}\gg \frac{\partial U(x_1,x_2)}{\partial x_2}$ ???and $dx_1/dx_2<0$??? the third party should reduce the incentives of firm $2$ to reduces its efforts and encourage firm $1$ to increase its efforts. While the total amount of effects does necessarily decrease, firm $1$'s efforts are seen as more important than firm $2$, creating a net benefit. If the effort do not change the firm's profit extraction, the latter case requires that $\frac{1}{\gamma_1}<g'(\theta^*)<0$.

In the smartphone sector ???does $x$ affect $\gamma$??? we observe: a small level of loyalty and a medium amount of asymmetry. Thus, supporting the weak firm will raise overall competition ???using which definition? same as above??? leading to a welfare increase. Similar, damaging the profit margin of the leader ???but $\gamma_{2}\leq \gamma_1$??? might lead to an overall increase in the efforts if the change in the number of indifferent customers is sufficiently big. Here moving closer to symmetry increases competition intensity.

???give more details for these examples: $d\gamma/dx >0$; what is effect of assumptions on $dx_1/dx_{2}$???

@@

???why repeat the expressions???

@@

???so what is the benchmark that readers expect; strategic substitutes???

@@

If customer loyalty is missing from the model, the only channel through which the efforts of the firms interact is the value extraction. If firm $2$ expands its market share at the expenses of firm $1$, firm $2$ has an additional incentive ???do you mean by ``additional'' that $dx_{1}/dx_{2} <0$? But in Nash equil. 2 takes $x_1$ as given??? to have higher efforts and firm $1$ to reduce its efforts. This fact does not depend on the market shares, so supporting the weak firm always reduces the efforts of the leader.

@@

In the next period, both firms exert less efforts as $g(\theta^*)$ goes down. However, the stronger firm ($1$) reacts more strongly to this, shifting market shares to firm $2$. Ironically, the market shares become more equal. It is important to note, that this builds on myopic firms.  
\section{Other aspects}
\begin{itemize}
	\item Market shares determine investments. Are returns convex or concave?
	\item Customers profit from investment. The large firm has more customers. If it invests more, it helps more people. ???socially optimal asymmetry???
	\item What is the social optimum?
	\item Switching costs?
\end{itemize}
